---
title: "HLCM.1 DAST speech intelligibility scores"
author: "Lars Bramsløw (LABW), Eriksholm Research Centre, Demant"
output:
  html_document:
    toc: true
    toc_float:
      toc_collapsed: true
    toc_depth: 3
    css: custom.css
  pdf_document:
    toc: true
    toc_depth: '3'
  word_document:
    toc: false
    toc_depth: '3'
header-includes: \usepackage{svg}
---

|Revision |Date | Author | Change  | 
|:----:|:----:|:----:|:----:|
|1.0| 24.02.2026 | LABW | Document created |
|1.1| 6.03.2026 | LABW | Analysis revised |
|1.2| 18.03.2026 | LABW | Special version without TP5 (anonymous) |
|1.3| 19.03.2026 | LABW | Full, normal analysis |


# 1 Read data and process

Note: all word keyword scores were rau-transformed in MATLAB.

```{r read, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

# TODO: 

library(readxl)
library(lmerTest)
library(ggplot2)
library(emmeans)

## read and convert data if need be
tdast <- read_excel("HLCM1_DAST.xlsx")
tdast$TPnr <- as.factor(tdast$TPnr)
tdast$Block <- as.factor(tdast$Block)
tdast$List <- as.factor(tdast$List)
tdast$SNR <- as.factor(tdast$SNR)
tdast$Talker <- as.factor(tdast$Talker)
tdast$Condition <- as.factor(tdast$Condition)

# remove unused columns
tdast$Mode <- NULL
tdast$`List type` <- NULL
tdast$`SRT in quiet` <- NULL

# remove training lists
tdast <- tdast[tdast$List != 11, ]

# remove TP5
# tdast <- tdast[tdast$TPnr != 5, ]


# show it
str(tdast)

### View(tdast)
```
Nr of TP read: `r length(levels(tdast$TPnr))`

# 2 Mixed model fitting

Do the mixed effects model fitting and show anova for fixed factors + ranova for random factors: 
```{r mmfit, echo=TRUE}
# Testing main effects and interactions of SNR and Condition
# Talker was found to have no effect 
# TP*Condition interaction was significant overall but no clear effects per individual
formula <- 'Score ~ 1 + SNR*Condition + (1 | TPnr)'   # full model

# lmerTest fitting the model
mdl <- lmerTest::lmer(formula, data = tdast)
anova(mdl)
ranova(mdl)
```
Significant effect of SNR.

No effect of Condition, neither SNR_Condition interaction. 

Large TP (random) effect, as often seen. 


Earlier exploratory tests:
Showed no relevant interaction TP_Condition.
Talker was found to have no effect. 
TP_Condition interaction was significant overall, but no clear effects per individual.

```{r summary, include=FALSE}
summary(mdl)

```


# 3 Estimated means

## 3.1 SNR

Large and expected effect of SNR:

```{r SNR, include=TRUE, message=FALSE, warning=FALSE}
#lsm <- ls_means(mdl)     #lmerTest means is not the best..
#plot(lsm, which = "scene")
#lsm <- ls_means(mdl, "scene")
em_SNR <- emmeans(mdl, "SNR")
#em_SNR

SNR_plot <- plot(em_SNR, horizontal = FALSE, xlab = 'Keyword score [%]')
SNR_plot = SNR_plot + theme_bw()
ggsave("SNR_plot.png", width = 100, height = 100, units = "mm")
plot(SNR_plot)

```

## 3.2 Processing condition

No significant main effect of condition:

```{r condition, message=FALSE, warning=FALSE}

em_Condition <- emmeans(mdl, "Condition")
#em_Condition

Condition_plot <- plot(em_Condition, horizontal = FALSE, xlab = 'Keyword score [%]')
Condition_plot = Condition_plot + theme_bw()
ggsave("Condition_plot.png", width = 100, height = 100, units = "mm")
plot(Condition_plot)

```

No significant interaction between SNR and condition. NAL-NL2 + NR looks best in all cases:

```{r SNR_condition, warning = FALSE, message = FALSE}
em_SNRCond <- emmeans(mdl, c("Condition", "SNR"))
#em_SNRCond

SNRCond_plot <- plot(em_SNRCond, horizontal = FALSE, xlab = 'Keyword score [%]')
SNRCond_plot = SNRCond_plot + theme_bw()
ggsave("SNRCond_plot.png", width = 100, height = 100, units = "mm")
plot(SNRCond_plot)
```

```{r TP}
# em_TP <- emmeans(mdl, 'TPnr')
# #em_TP
# 
# TP_plot <- plot(em_TP, horizontal = FALSE, xlab = 'Keyword score [%]')
# TP_plot = TP_plot + theme_bw() + theme(axis.text.x = element_blank())
# ggsave("TP_plot.png", width = 100, height = 100, units = "mm")
# plot(TP_plot)

# Large TP effect - typical in HI users


```


```{r TP_Condition}
# em_TP_Cond <- emmeans(mdl, c("Condition", "TPnr"))
# #em_TP_Cond
# 
# TP_Cond_plot <- plot(em_TP_Cond, horizontal = FALSE, xlab = 'Keyword score [%]')
# TP_Cond_plot = TP_Cond_plot + theme_bw() + 
# theme(
#   axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)
# )
# 
# ggsave("TP_Cond_plot.png", width = 400, height = 200, units = "mm")
# plot(TP_Cond_plot)

```



```{r}
# 3way interaction
# plot(lsm, which = "space:pos:attribut")
```


```{r}
# res <- plot(lsm, mult = FALSE)
```


# 4 Session info

```{r versions, echo=FALSE}
sessionInfo()
```